---
title: "ANGSD_HW_Week_10"
author: "Paul Gaudin"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Exercise 1

#####What do the columns represent? What do the rows represent? 

The column names are the barcodes for each cell, the row names are Gene names, Column x Gene is the gene count for each of the cells with a specified barcode. 

#####Create a SingleCellExperiment object.

```{r message=FALSE}
folder <- "./"

readcounts <- read.table(paste0(folder, "WT-1.dge.txt"),header=TRUE, row.names=1)

counts_matrix <- as.matrix(readcounts)

library(SingleCellExperiment)

sce <- SingleCellExperiment(assays = list(counts = counts_matrix, logcounts = log10(counts_matrix)))
sce
```

#####Show a snippet of the count matrix that is now part of the SCE object.
```{r message=FALSE}
counts(sce)[1:10,1:5]
```

#####Calculate the numbers of reads (~ sequencing depths) for the first five cells, i.e. you should have 5 values in the end.
```{r message=FALSE}
for (i in 1:5){
  print(colnames(counts(sce))[i]) ##barcode
  print(sum(counts(sce)[,i])) ##sequencing depth
}
```

#####How many genes have non-zero counts in the first five cells?
```{r message=FALSE}
genes_with_non_zero_counts <- 0
counts_matrix = counts(sce)
for (i in 1:nrow(counts_matrix)){
  if(sum(counts_matrix[i,1:5]) > 0){
    genes_with_non_zero_counts <- genes_with_non_zero_counts + 1
  }
}

genes_with_non_zero_counts
```

#####If you were to change the row- and column-names of the original matrix, how could you keep track of the original names? 

You could create an object that has keys that are the new row or column names, and values that are the original row and column names. 

#####Following the simpleSingleCell workflow, generate histograms or density plots of (a) the total number of UMI counts per cell and (b) total number of expressed genes. (1pt plus 1pt extra-credit if you generate the plots with ggplot2).

```{r message=FALSE}

#calculateQCMetrics(sce)

total_expressed_genes = c()
total_umi_count = c();
for (i in 1:ncol(counts_matrix)){
  total_expressed_genes = c(total_expressed_genes,sum(counts_matrix[,i] > 0))
}

total_UMI_count = c();
for (i in 1:ncol(counts_matrix)){
  total_umi_count = c(total_umi_count,sum(counts_matrix[,i]))
}

hist(log10(total_expressed_genes), breaks=30, col="grey80",
    xlab="Log10-total expressed genes")

hist(log10(total_umi_count), breaks=30, col="grey80",
    xlab="Log10-total UMI count")

```
